### 2024-03-27 docs/projects/geography/chicago-bikes/index.qmd
Chicago bikes - rerendered

---
<style></style><style>.printedClojure .sourceCode {
  background-color: transparent;
  border-style: none;
}
</style><style>.clay-limit-image-width .clay-image {max-width: 100%}
</style>
<script src="index_files/md-default43.js" type="text/javascript"></script><script src="index_files/md-default44.js" type="text/javascript"></script><script src="index_files/vega45.js" type="text/javascript"></script><script src="index_files/vega46.js" type="text/javascript"></script><script src="index_files/vega47.js" type="text/javascript"></script>

@@

::: {.sourceClojure}
```clojure
(defn as-geo [vega-lite-spec]
  (-> vega-lite-spec
      (update :encoding update-keys (fn [k]
                                      (get {:x :latitude
                                            :y :longitude
                                            :x2 :latitude2
                                            :y2 :longitude2}
                                           k k)))
      (assoc :projection {:type :mercator}
             ;; rendering as PNG this time
             ;; to make this page more lightweight
             :usermeta {:embedOptions {:renderer :png}})))
```
:::

@@

::: {.sourceClojure}
```clojure
(delay
  (-> processed-trips
      (tc/random 10000 {:seed 1})
      (hanami/plot ht/rule-chart
                   {:X :start_lat
                    :Y :start_lng
                    :X2 :end_lat
                    :Y2 :end_lng
                    :OPACITY 0.1})
      as-geo))
```
:::

@@

::: {.sourceClojure}
```clojure
(delay
  (-> processed-trips
      (tc/random 10000 {:seed 1})
      (hanami/plot ht/layer-chart
                   {:TITLE "Chicago bike trips"
                    :LAYER [{:data {:url "notebooks/data/chicago.geojson"
                                    :format {:type "topojson"}}
                             :mark {:type "geoshape"
                                    :filled false
                                    :clip true
                                    :opacity 0.5}}
                            (as-geo
                             (hanami/plot nil ; use the global dataset for this layer
                                          ht/rule-chart
                                          {:X :start_lat
                                           :Y :start_lng
                                           :X2 :end_lat
                                           :Y2 :end_lng
                                           :OPACITY 0.1}))]})))
```
:::

```{=html}
<div><script>vegaEmbed(document.currentScript.parentElement, {"usermeta":{"embedOptions":{"renderer":"svg"}},"width":400,"background":"floralwhite","title":{"text":"Chicago bike trips"},"layer":[{"data":{"url":"notebooks\/data\/chicago.geojson","format":{"type":"topojson"}},"mark":{"type":"geoshape","filled":false,"clip":true,"opacity":0.5}},{"encoding":{"longitude":{"field":"start_lng","type":"quantitative"},"opacity":{"value":0.1},"latitude":{"field":"start_lat","type":"quantitative"},"longitude2":{"field":"end_lng","type":"quantitative"},"latitude2":{"field":"end_lat","type":"quantitative"}},"usermeta":{"embedOptions":{"renderer":"png"}},"mark":"rule","width":400,"background":"floralwhite","projection":{"type":"mercator"},"height":300}],"height":300,"data":{"url":"index_files\/2.csv","format":{"type":"csv"}}});</script></div>
```

@@

::: {.sourceClojure}
```clojure
(delay
  (-> processed-trips
      (tc/add-column :cluster (:clustering clustering))
      (tc/group-by [:cluster])
      ;; sort the groups in ascending order by their size,
      ;; and take the first 20 (which are the largest ones)
      ;; (we use `without-grouping->` to order and take
      ;; across groups rather tan within groups):
      (tc/without-grouping->
       (tc/order-by (fn [ds]
                      (-> ds :data tc/row-count))
                    :desc)
       (tc/head 20))
      (tc/aggregate {:n tc/row-count
                     :map (fn [trips]
                            [(-> trips
                                 (hanami/plot ht/layer-chart
                                              {:TITLE "Chicago bike trips"
                                               :LAYER [{:data {:url "notebooks/data/chicago.geojson"
                                                               :format {:type "topojson"}}
                                                        :mark {:type "geoshape"
                                                               :filled false
                                                               :clip true
                                                               :opacity 0.3}}
                                                       (as-geo
                                                        (hanami/plot nil
                                                                     ht/rule-chart
                                                                     {:X :start_lat
                                                                      :Y :start_lng
                                                                      :X2 :end_lat
                                                                      :Y2 :end_lng
                                                                      :OPACITY 0.01}))
                                                       (as-geo
                                                        (hanami/plot nil
                                                                     ht/point-chart
                                                                     {:X :start_lat
                                                                      :Y :start_lng
                                                                      :MCOLOR "purple"
                                                                      :OPACITY 0.01}))
                                                       (as-geo
                                                        (hanami/plot nil
                                                                     ht/point-chart
                                                                     {:X :end_lat
                                                                      :Y :end_lng
                                                                      :MCOLOR "green"
                                                                      :OPACITY 0.01}))]}))])
                     :hours (fn [trips]
                              [(hour-counts-plot trips)])})
      kind/table))
```
:::

>>>>

---
<style></style><style>.printedClojure .sourceCode {
  background-color: transparent;
  border-style: none;
}
</style><style>.clay-limit-image-width .clay-image {max-width: 100%}
</style>
<script src="index_files/md-default23.js" type="text/javascript"></script><script src="index_files/md-default24.js" type="text/javascript"></script><script src="index_files/vega25.js" type="text/javascript"></script><script src="index_files/vega26.js" type="text/javascript"></script><script src="index_files/vega27.js" type="text/javascript"></script>

@@

::: {.sourceClojure}
```clojure
(defn as-geo [vega-lite-spec]
  (-> vega-lite-spec
      (update :encoding update-keys (fn [k]
                                      (get {:x :latitude
                                            :y :longitude
                                            :x2 :latitude2
                                            :y2 :longitude2}
                                           k k)))
      (assoc :projection {:type :mercator})))
```
:::

We will find it useful to render some Vega-Lite plots
as PNG, rather than SVG. This will meke the whole page more lightweight.

::: {.sourceClojure}
```clojure
(defn as-png [vega-lite-spec]
  (-> vega-lite-spec
      (assoc :usermeta {:embedOptions {:renderer :png}})))
```
:::

@@

::: {.sourceClojure}
```clojure
(delay
  (-> processed-trips
      (tc/random 10000 {:seed 1})
      (hanami/plot ht/rule-chart
                   {:X :start_lat
                    :Y :start_lng
                    :X2 :end_lat
                    :Y2 :end_lng
                    :OPACITY 0.1})
      as-geo
      as-png))
```
:::

@@

::: {.sourceClojure}
```clojure
(delay
  (-> processed-trips
      (tc/random 10000 {:seed 1})
      (hanami/plot ht/layer-chart
                   {:TITLE "Chicago bike trips"
                    :LAYER [{:data {:url "notebooks/data/chicago.geojson"
                                    :format {:type "topojson"}}
                             :mark {:type "geoshape"
                                    :filled false
                                    :clip true
                                    :opacity 0.5}}
                            (as-geo
                             (hanami/plot nil ; use the global dataset for this layer
                                          ht/rule-chart
                                          {:X :start_lat
                                           :Y :start_lng
                                           :X2 :end_lat
                                           :Y2 :end_lng
                                           :OPACITY 0.1}))]})
      as-png))
```
:::

```{=html}
<div><script>vegaEmbed(document.currentScript.parentElement, {"usermeta":{"embedOptions":{"renderer":"png"}},"width":400,"background":"floralwhite","title":{"text":"Chicago bike trips"},"layer":[{"data":{"url":"notebooks\/data\/chicago.geojson","format":{"type":"topojson"}},"mark":{"type":"geoshape","filled":false,"clip":true,"opacity":0.5}},{"encoding":{"longitude":{"field":"start_lng","type":"quantitative"},"opacity":{"value":0.1},"latitude":{"field":"start_lat","type":"quantitative"},"longitude2":{"field":"end_lng","type":"quantitative"},"latitude2":{"field":"end_lat","type":"quantitative"}},"usermeta":{"embedOptions":{"renderer":"svg"}},"mark":"rule","width":400,"background":"floralwhite","projection":{"type":"mercator"},"height":300}],"height":300,"data":{"url":"index_files\/2.csv","format":{"type":"csv"}}});</script></div>
```

@@

::: {.sourceClojure}
```clojure
(delay
  (-> processed-trips
      (tc/add-column :cluster (:clustering clustering))
      (tc/group-by [:cluster])
      ;; sort the groups in ascending order by their size,
      ;; and take the first few (which are the largest ones)
      ;; (we use `without-grouping->` to order and take
      ;; across groups rather tan within groups):
      (tc/without-grouping->
       (tc/order-by (fn [ds]
                      (-> ds :data tc/row-count))
                    :desc)
       (tc/head 10))
      (tc/aggregate {:n tc/row-count
                     :map (fn [trips]
                            [(-> trips
                                 (hanami/plot ht/layer-chart
                                              {:TITLE "Chicago bike trips"
                                               :LAYER [{:data {:url "notebooks/data/chicago.geojson"
                                                               :format {:type "topojson"}}
                                                        :mark {:type "geoshape"
                                                               :filled false
                                                               :clip true
                                                               :opacity 0.3}}
                                                       (as-geo
                                                        (hanami/plot nil
                                                                     ht/rule-chart
                                                                     {:X :start_lat
                                                                      :Y :start_lng
                                                                      :X2 :end_lat
                                                                      :Y2 :end_lng
                                                                      :OPACITY 0.01}))
                                                       (as-geo
                                                        (hanami/plot nil
                                                                     ht/point-chart
                                                                     {:X :start_lat
                                                                      :Y :start_lng
                                                                      :MCOLOR "purple"
                                                                      :OPACITY 0.01}))
                                                       (as-geo
                                                        (hanami/plot nil
                                                                     ht/point-chart
                                                                     {:X :end_lat
                                                                      :Y :end_lng
                                                                      :MCOLOR "green"
                                                                      :OPACITY 0.01}))]})
                                 as-png)])
                     :hours (fn [trips]
                              [(as-png
                                (hour-counts-plot trips))])})
      kind/table))
```
:::
